---
title: "p8105_hw2_zdz2101"
author: "Zelos Zhu"
date: "9/25/2018"
output: github_document
---

#Problem 1
```{r NYC Transit, message = FALSE}
library(tidyverse)
nyc_transit_df <- 
  read_csv("Data Files/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(., line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6,
                route7, route8, route9, route10, route11, entrance_type, entry, vending, ada) %>% 
  mutate(entry = recode(entry, "YES" = 1, "NO" = 0))

#Distinct Stations
unique(apply(nyc_transit_df[,names(nyc_transit_df)[c(1,2,5:15)]], 1, paste , collapse = "/" ))

#ADA compliant
table(nyc_transit_df$ada)

#Vending and entrances
table(nyc_transit_df$entry, nyc_transit_df$vending)
```

#Problem 2
This problem uses the Mr. Trash Wheel dataset, available as an Excel file on the course website.

Read and clean the Mr. Trash Wheel sheet:
-specify the sheet in the Excel file and to omit columns containing notes (using the range argument and cell_cols() function) 
-use reasonable variable names 
-omit rows that do not include dumpster-specific data 
-rounds the number of sports balls to the nearest integer and converts the result to an integer variable (using  as.integer).

Read and clean precipitation data for 2016 and 2017. For each, omit rows without precipitation data and add a variable year. Next, combine datasets and convert month to a character variable (the variable month.name is built into R and should be useful).

Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2017? What was the median number of sports balls in a dumpster in 2016?

```{r, Mr. Trash Wheel}
library(readxl)
mr_trash_wheel <- read_excel("Data Files/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>%
  janitor::clean_names() %>%
  filter(!grepl("Total", month)) %>%
  mutate(sports_balls = as.integer(round(sports_balls)))

precip_2016_df <- read_excel("Data Files/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2016 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = rep(2016, nrow(.)))

precip_2017_df <- read_excel("Data Files/HealthyHarborWaterWheelTotals2017-9-26.xlsx", sheet = "2017 Precipitation", range = "A2:B10") %>%
  janitor::clean_names() %>%
  mutate(year = rep(2017, nrow(.)))

precip_df <- rbind(precip_2016_df, precip_2017_df) %>% 
  mutate(month = month.name[month])
```

#Problem 3
This problem uses the BRFSS data. DO NOT include this dataset in your local data directory; instead, load the data from the  p8105.datasets package.

For this question:
-format the data to use appropriate variable names;
-focus on the “Overall Health” topic
-exclude variables for class, topic, question, sample size, and everything from lower confidence limit to GeoLocation
-structure data so that responses (excellent to poor) are variables taking the value of Data_value
-create a new variable showing the proportion of responses that were “Excellent” or “Very Good”

Using this dataset, do or answer the following:

-How many unique locations are included in the dataset? Is every state represented? What state is observed the most?
-In 2002, what is the median of the “Excellent” response value?
-Make a histogram of “Excellent” response values in the year 2002.
-Make a scatterplot showing the proportion of “Excellent” response values in New York County and Queens County (both in NY State) in each year from 2002 to 2010.
```{r}
devtools::install_github("p8105/p8105.datasets")
library(p8105.datasets)
data("brfss_smart2010")
head(brfss_smart2010)
```